Copykat CNV results exploration for `r params$sample_id`

### 2024-10-15 analyses/cell-type-wilms-tumor-06/notebook_template/05_copykat_exploration.Rmd
Start exploring copykat results

---
title: "Copykat CNV results exploration for `r params$sample_id`"
author: "Maud PLASCHKA"
date: '2024-08-13'
params:
  scpca_project_id: "SCPCP000006"
  sample_id: "SCPCS000194"
  distance: "spearman"
  seed: 12345
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

@@

subdiagnosis <- readr::read_tsv(
  file.path("..", "..", "..", "data", "current", params$scpca_project_id, "single_cell_metadata.tsv"),
  show_col_types = FALSE
  ) |>
  dplyr::filter(scpca_sample_id == params$sample_id) |>
  dplyr::pull(subdiagnosis)

@@

`CopyKAT` was run using the `copyKAT.R` script with and without a normal reference, using `r params$distance` to calculate distance in `copyKAT`.

@@

In this notebook, we are working with the Wilms tumor sample defined in `r params$sample_id` from the Wilms tumor dataset `r params$scpca_project_id`.
We work with the pre-processed and labeled `Seurat` object that is the output of `02b_label-transfer_fetal_kidney_reference_Stewart.Rmd` saved in the `results` directory.

@@

result_dir <- file.path(module_base, "results", params$sample_id)

# predictions files 
predictions_file <- glue::glue("{params$sample_id}_copykat_prediction.txt")

predictions_paths <- c(
  no_ref = file.path(result_dir, "05_copyKAT", "noref", params$distance, predictions_file),
  with_ref = file.path(result_dir, "05_copyKAT", "ref", params$distance, predictions_file)
)

# full results gene by cell 
full_ck_result_file <- glue::glue("{params$sample_id}_copykat_CNA_results.txt")

full_ck_result_paths <- c(
  no_ref = file.path(result_dir, "05_copyKAT", "noref", params$distance, full_ck_result_file),
  with_ref = file.path(result_dir, "05_copyKAT", "ref", params$distance, full_ck_result_file)
)

@@

![](`r file.path(result_dir, "copyKAT", "noref", params$distance, glue::glue("{params$sample_id}_copykat_heatmap.jpeg"))`)

@@

![](`r file.path(result_dir, "copyKAT", "ref", params$distance, glue::glue("{params$sample_id}_copykat_heatmap.jpeg"))`)

@@

# join with cnv info
cnv_df <- cnv_df |>
  dplyr::left_join(full_cnv_df, by = c("barcodes", "reference_used")) |>
  dplyr::filter(!is.na(chrom))
```

@@

>>>>

---
title: "Copykat CNV results exploration for `r params$sample_id`"
author: "Maud PLASCHKA"
date: '2024-08-13'
params:
  sample_id: "SCPCS000194"
  seed: 12345
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

@@

subdiagnosis <- readr::read_tsv(
  file.path("..", "..", "..", "data", "current", "SCPCP000006", "single_cell_metadata.tsv"),
  show_col_types = FALSE
) |>
  dplyr::filter(scpca_sample_id == params$sample_id) |>
  dplyr::pull(subdiagnosis)

@@

`CopyKAT` was run using the `copyKAT.R` script with and without a normal reference, using either an euclidean or statistical (spearman) method to calculate distance in `copyKAT`.

@@

In this notebook, we are working with the Wilms tumor sample defined in `r params$sample_id` from the Wilms tumor dataset "SCPCP000006".
We work with the pre-processed and labeled `Seurat` object that is the output of `02b_label-transfer_fetal_kidney_reference_Stewart.Rmd` saved in the `results` directory.

@@

result_dir <- file.path(module_base, "results", params$sample_id)
predictions_paths <- list()
full_ck_result_paths <- list()
for(ref_value in c("ref", "noref")){
  for(distance_value in c("euclidean", "spearman")){
    predictions_file <- glue::glue("05_copykat_", {params$sample_id}, "_",ref_value,"_distance-", distance_value, "_copykat_prediction.txt")
    full_ck_result_file <- glue::glue("05_copykat_", {params$sample_id}, "_",ref_value,"_distance-", distance_value, "_copykat_CNA_results.txt")

    predictions_paths[[glue::glue(ref_value, "_", distance_value)]] <- file.path(result_dir, "05_copyKAT", ref_value, distance_value, predictions_file)
    full_ck_result_paths[[glue::glue(ref_value, "_", distance_value)]] <- file.path(result_dir, "05_copyKAT", ref_value, distance_value, full_ck_result_file)
  }
}
```

@@

##### Euclidean distance

![](`r file.path(result_dir, "05_copyKAT", "noref", "euclidean", glue::glue("05_copykat_", params$sample_id, "_noref_distance-euclidean_copykat_heatmap.png"))`)

##### Spearman distance

![](`r file.path(result_dir, "05_copyKAT", "noref", "spearman", glue::glue("05_copykat_", params$sample_id, "_noref_distance-spearman_copykat_heatmap.png"))`)

@@

##### Euclidean distance

![](`r file.path(result_dir, "05_copyKAT", "ref", "euclidean", glue::glue("05_copykat_", params$sample_id, "_ref_distance-euclidean_copykat_heatmap.png"))`)

##### Spearman distance

![](`r file.path(result_dir, "05_copyKAT", "ref", "spearman", glue::glue("05_copykat_", params$sample_id, "_ref_distance-spearman_copykat_heatmap.png"))`)

@@

# join with cnv info
cnv_df <- cnv_df |>
   dplyr::left_join(full_cnv_df, by = c("barcodes", "reference_used")) |>
   dplyr::filter(!is.na(chrom))

@@

From the heatmap of CNV and the mean CNV detection plots, it is difficult to see any CNV pattern that drive the identification of aneuploid cells. 
The assignment of the aneuploidy/diploidy value might relies on very few CNV and/or an arbitrary threshold. 
This might be why the assignement of aneuploidy/diploidy values differs between condition (and between runs!!).
